---
title: "'Ground truth': phrase boundaries"
author: "Riesa Cassano"
date: "3/28/2022"
output: html_document
---

```{r libraries, include=FALSE}
library(tidyverse)
library(magrittr)
```

Here are the 'ground truth' labels that indicate phrase structure in the stimuli. From these labels, we generate and save a data frame that is in the same format as `responses_binarized_long.csv` for use with that data.

#### Description
Actual phrase boundaries are extracted from the Intact and 8B conditions. Since these conditions have intact phrases, these phrase boundaries are meaningful in that they are descriptive of actual structure in the stimuli. 

Pseudo-phrase boundaries are constructed for the 2B and 1B conditions. This is done in two ways. First, you could simply mark every eight bars, approximately the length of actual phrases in Intact and 8B. These conditions do not have intact phrases, so these pseudo-phrase boundaries don't reflect structure in the stimuli. We expect that these pseudo-phrase boundaries will have nothing to do with the boundaries that the subjects actually perceive. However, it is possible that the subjects have a 'natural' response rate that is close to 8-bar phrases - using these pseudo-phrase boundaries to predict subject responses might pick up on some regularity in the response data. Second, you could mark the last bar of the phrases in intact and see where that particular bar ends up in the scramble. If the last bar on its own is convincing as a phrase ending (and if the preceding context is plausible), subjects might respond.

#### Time points
Phrases in **Intact** (original, long stimulus) begin at:
1,19,35,51,67,79,95,111,126,143,159,175,193,209,225

The original stimulus is split into thirds:

* intact_1 (seconds 1-78) contains boundaries at 19,35,51,67

* intact_2 (seconds 79-158) contains boundaries at 17,33,49,65

* intact_3 (seconds 159-240) contains boundaries at 17,35,51,67


Phrases in **8B** (original, long stimulus) begin at:
1,19,35,51,67,85,101,117,133,149,165,181,197,213,225

Again, the original stimulus is split into thirds:

* 8B_1 (seconds 0-84) contains boundaries at 19,35,51,67

* 8B_2 (seconds 85-164) contains boundaries at 17,33,49,65

* 8B_3 (seconds 165-240) contains boundaries at 17,33,49,61


*Periodic:* Pseudo-phrase boundaries in **2B** and **1B** are identical. They end at: `16*x` since each 8-bar phrase is 16 seconds long (each bar is 2 beats, each beat is one second). This places more importance on the expected regularity of music.

*Tracked:* Phrase endings in **2B** and **1B** are tracked from **Intact** through the scramble. This destroys the periodic structure. Instead, this places more importance on the features of the ending bar itself.

* 2B_1 (seconds 0-80) contains tracked boundaries at 2,10,56

* 2B_2 (seconds 80-160) contains tracked boundaries at 2,6,28,38,46,64

* 2B_3 (seconds 160-240) contains tracked boundaries at 18,30,58,62,66,78

* 1B_1 (seconds 0-80) contains tracked boundaries at 10,22,42,70,74

* 1B_2 (seconds 80-160) contains tracked boundaries at 6,20,44,50

* 1B_3 (seconds 160-240) contains tracked boundaries at 4,16,20,34,36,80


#### Timepoints in a usable format (binary vector)
In one generated csv (`ground_truth_phrase_boundaries_periodic.csv`), phrase boundaries (intact, 8B) and pseudo-phrase boundaries (2B, 1B) are marked at the end of the phrase, with a 1 in the last second of the phrase (0 elsewhere).

```{r generate periodic, echo=FALSE}
scram_cond <- c('intact','8B','2B','1B')
stim_num <- c(1,2,3)
# each stimulus gets a row in the phrase boundaries (pb) dataframe
pb <- crossing(scram_cond, stim_num)

# for the Intact and 8B phrases, we include the end of the last phrase of the stimuli, to make it more accurate with the padding at the end

# last second of each pseudo-phrase in 2B and 1B stimuli
pseudo_pb <- c(16,32,48,64,80)

all_pb <- list(pseudo_pb, # 1B 1
               pseudo_pb, # 1B 2
               pseudo_pb, # 1B 3
               pseudo_pb, # 2B 1
               pseudo_pb, # 2B 2
               pseudo_pb, # 2B 3
               c(18,34,50,66,84), # 8B 1 19,35,51,67
               c(16,32,48,64,80), # 8B 2 17,33,49,65 
               c(16,32,48,60,76), # 8B 3 17,33,49,61 
               c(18,34,50,66,78), # Intact 1 19,35,51,67
               c(16,32,48,64,80), # Intact 2 17,33,49,65
               c(16,34,50,66,82)) # Intact 3 17,35,51,67

pb$pb_vals <- all_pb

# The longest stimulus is 1:25, so the binary vectors are at most 85 seconds long.
response_vec_length <- 85
# There are two "identification" columns (scram_cond, stim_num) and then the data (pb_vals)
id_cols <- ncol(pb)

# add the columns of zeros
seconds <- as.character(0:response_vec_length)
pb[seconds] <- 0

# use pb_vals to determine which seconds have a response in them
for (i in 1:nrow(pb)){
  
  # grab the response values for this row and make them ints
  these_vals <- as.integer(unlist(pb$pb_vals[i]))
  
  # use those values to index in the columns we created
  for (j in these_vals){
    pb[[j+id_cols+1]][i] <- 1 # id_cols+1 is for '0'
  }
  
}

pb$pb_vals <- NULL

# convert from wide to long
pb %>%
  pivot_longer(-c(scram_cond,stim_num), 
               names_to = 'time', values_to = 'response') %>%
  #write_csv('data/ground_truth_phrase_boundaries_periodic.csv')
  print()

```

In the other generated csv (`ground_truth_phrase_boundaries_tracked.csv`), phrase boundaries for intact and 8B are the same. For 2B and 1B the actual phrase endings are tracked through the scramble. 

```{r generate tracking, echo=FALSE}
# each stimulus gets a row in the phrase boundaries (pb) dataframe
pb <- crossing(scram_cond, stim_num)

# for the Intact and 8B phrases, we include the end of the last phrase of the stimuli, to make it more accurate with the padding at the end

all_pb <- list(c(10,22,42,70,74), # 1B 1
               c(6,20,44,50), # 1B 2
               c(4,16,20,34,36,80), # 1B 3
               c(2,10,56), # 2B 1
               c(2,6,28,38,46,64), # 2B 2
               c(18,30,58,62,66,78), # 2B 3
               c(18,34,50,66,84), # 8B 1 19,35,51,67
               c(16,32,48,64,80), # 8B 2 17,33,49,65 
               c(16,32,48,60,76), # 8B 3 17,33,49,61 
               c(18,34,50,66,78), # Intact 1 19,35,51,67
               c(16,32,48,64,80), # Intact 2 17,33,49,65
               c(16,34,50,66,82)) # Intact 3 17,35,51,67

pb$pb_vals <- all_pb

# The longest stimulus is 1:25, so the binary vectors are at most 85 seconds long.
response_vec_length <- 85
# There are two "identification" columns (scram_cond, stim_num) and then the data (pb_vals)
id_cols <- ncol(pb)

# add the columns of zeros
seconds <- as.character(0:response_vec_length)
pb[seconds] <- 0

# use pb_vals to determine which seconds have a response in them
for (i in 1:nrow(pb)){
  
  # grab the response values for this row and make them ints
  these_vals <- as.integer(unlist(pb$pb_vals[i]))
  
  # use those values to index in the columns we created
  for (j in these_vals){
    pb[[j+id_cols+1]][i] <- 1 # id_cols+1 is for '0'
  }
  
}

pb$pb_vals <- NULL

# convert from wide to long
pb %>%
  pivot_longer(-c(scram_cond,stim_num), 
               names_to = 'time', values_to = 'response') %>%
  #write_csv('data/ground_truth_phrase_boundaries_tracked.csv')
  print()

```